\section{Conclusion}

Satellite embeddings are becoming a practical abstraction layer for Earth observation. They connect large archives, foundation-model training, precomputed products, and lightweight downstream analysis, but the literature still tends to describe them indirectly through model families or benchmark datasets. This review argues for a more direct framing: the reusable representation itself should be the unit of organization. From that perspective, precomputed products, masked autoencoders, vision-language encoders, multimodal generalists, and domain-specialized models can be compared as different ways of producing transferable geospatial features under different sensing assumptions.

The main contribution of the review is therefore twofold. First, it provides an embedding-centered taxonomy and model dossier for \modelcount{} representative public models, making sensor support, temporal semantics, output form, dimensionality, and deployment mode visible alongside architecture and pretraining objective. Second, it links that taxonomy to a compact Wuhan benchmark that evaluates frozen embeddings under shared geographic, label, and probe conditions. The results show that model choice depends on the target property being tested: semantic separability, continuous land-cover mixture, water thresholdability, and dense spatial separability emphasize different strengths and weaknesses. They also show that scale mismatch is not a minor implementation detail, but a central factor in whether an embedding is meaningful for local mapping.

The broader lesson is that satellite embeddings should be selected and evaluated as geospatial representations, not only as checkpoints. A useful review must therefore connect literature synthesis with data regimes, application requirements, operational constraints, and empirical diagnostics. The current paper is a step in that direction: it keeps the broad survey structure of recent remote sensing reviews, but adds an embedding-specific benchmark lens that makes representation behavior visible in one controlled ROI. Future extensions should scale this evidence across regions, years, and physically grounded regression targets while preserving the same principle: compare embeddings in the form practitioners actually use them.
